Current status and propsects of R-packages for the design of experiments

### 2021-02-09 paper.Rmd
high level structure of the paper + some cosmetics for plots changed

---
title: Current status and propsects of R-packages for the design of experiments
author:
  # see ?rjournal_article for more information
  - name: Emi Tanaka
    affiliation: Monash University
    address:
    - Monash University
    - Clayton campus, VIC 3800, Australia
    url: http://emitanaka.org/
    orcid: 0000-0002-1455-259X
    email:  emi.tanaka@monash.edu
abstract: >
  An abstract of less than 150 words.
preamble: |
  % Any extra LaTeX you need in the preamble

@@

@@

knitr::opts_chunk$set(echo = FALSE, 
                      cache = TRUE,
                      cache.path = "cache/",
                      fig.align='center', 
                      fig.pos='htbp')

myggtheme <- 
  theme(panel.background = element_rect(fill = NA),
        panel.grid = element_line(color = "#f6e5ee"),
        axis.text = element_text(color = "#79003e"),
        axis.line = element_line(color = "#79003e", size = 0.7),
        axis.ticks.length = unit(1.4, "mm"),
        axis.ticks = element_line(color = "#79003e", size = 0.7),
        axis.title = element_text(color = "#79003e", face = "bold"),
        strip.background = element_rect(color = "#79003e",
                                        fill = "#AD0059"),
        strip.text = element_text(color = "white"),
        plot.title.position = "plot",
        plot.title = element_text(color = "#79003e", face = "bold")) 
```

@@

```{r bigram-title}
bigram_tab(doe_db, Title) %>% 
  filter(n > 3) %>% 
  mutate(word = paste(word1, word2)) %>% 
  select(word, n) %>% 
  kbl(caption = "The bigram of the R-package titles as provided in the DESCRIPTION file in CRAN.", 
      col.names = c("Bigram", "Count")) 
```

```{r bigram-desc}
bigram_tab(doe_db, Description) %>% 
  filter(n > 4) %>% 
  mutate(word = paste(word1, word2)) %>% 
  select(word, n) %>% 
  kbl(caption = "The bigram of the R-package descriptions as provided in the DESCRIPTION file in CRAN.", 
      col.names = c("Bigram", "Count"), 
      booktabs = TRUE) 
```

@@

```{r cran-data}
end <- Sys.Date() - 2 # usually 1-2 days are not available yet
start <- end - years(5) + days(2)
dldat <- cran_downloads(doe_pkgs, from = start, to = end)
```

```{r download-hist}
dldat %>% 
  group_by(package) %>% 
  summarise(total = sum(count)) %>%
  ggplot(aes(total)) + 
  geom_histogram(color = "white", fill = "#AD0059") + 
  scale_x_log10(label = comma) + 
  myggtheme + 
  labs(x = glue("Total download counts from {start} to {end}"),
       y = "Number of packages") +
  scale_y_continuous(expand = c(0, 0))
```

```{r top5}
ntop <- 5

@@

```{r download-barplot}
top5sum_df %>% 
  mutate(package = fct_reorder(package, total)) %>% 
  ggplot(aes(total, package)) +
  geom_col(aes(fill = package)) +
  labs(x = glue("Total downloads from {start} to {end}"),
       y = "Package") + 
  scale_x_continuous(labels = comma, expand = c(0, 0)) +
  myggtheme + 
  scale_fill_discrete_qualitative(rev = TRUE) + 
  guides(fill = FALSE)
```

```{r pkg-updates, cache = TRUE, dependson="doe-cran"}
pkg_url <- "https://cran.r-project.org/web/packages/{pkg}/index.html"
pkg_archive <- "https://cran.r-project.org/src/contrib/Archive/{pkg}/"
pkg_updates <- map(top5, function(pkg) {
  last_update <- read_html(glue(pkg_url)) %>% 
    html_table() %>% 
    .[[1]] %>% 
    filter(X1=="Published:") %>% 
    pull(X2) %>% 
    ymd()

@@

```{r download-timeplot, dependson="top5"}
ggplot(top5_df, aes(date, count, color = package)) +
  # add shadow lines
  geom_line(data = rename(top5_df, package2 = package), 
            color = "gray", aes(group = package2)) +
  # add date when package was updated
  geom_vline(data = updates, aes(xintercept = update),
             linetype = "dashed", color = "#79003e") + 
  # the trend line
  geom_line() +
  scale_y_log10() +
  facet_grid(package ~ .) + 
  labs(title = glue("Top 5 downloaded DoE packages from {start} to {end}")) + 
  scale_color_discrete_qualitative() +
  guides(color = FALSE) +
  myggtheme
```

>>>>

---
title: Current status and propsects of R-packages for the design of experiments
author:
  # see ?rjournal_article for more information
  - name: Emi Tanaka
    affiliation: Monash University
    address:
    - Monash University
    - Clayton campus, VIC 3800, Australia
    url: http://emitanaka.org/
    orcid: 0000-0002-1455-259X
    email:  emi.tanaka@monash.edu
abstract: >
  The critical role of data collection is well captured in the expression ``garbage in, garbage out'' -- in other words, if the collected data is rubbish then no analysis, however complex it may be, can make something out of it. The gold standard for data collection is through well-designed experiments.  Re-running an experiment is generally expensive and in some cases difficult due to limited resources, The design of experiment is important in getting a good value out of the collected data. In this article I describe the current state of the R-packages for the design of experiments through textual analysis and download trends. I discuss also the software design of widely utilised R packages in the field and conclude with discussion of some future prospects for the field.
preamble: |
  % Any extra LaTeX you need in the preamble

@@

# Introduction

# Explorative Analysis

## What are the common types of experimental designs?

## How do packages interplay with each other?

## Which packages are widely utilised?

# Software Design

# Propsects and Discussion

\newpage

@@

knitr::opts_chunk$set(echo = FALSE, 
                      cache = TRUE,
                      cache.path = "cache/",
                      fig.align='center', 
                      fig.pos='htbp', 
                      fig.width = 6,
                      message = FALSE,
                      warning = FALSE)

theme_set(
  theme(panel.background = element_rect(fill = NA),
        panel.grid = element_line(color = "lightgray"),
        axis.text = element_text(color = "black"),
        axis.line = element_line(color = "black", size = 0.7),
        axis.ticks.length = unit(1.4, "mm"),
        axis.ticks = element_line(color = "black", size = 0.7),
        axis.title = element_text(color = "black", face = "bold"),
        strip.background = element_rect(color = "black",
                                        fill = "black"),
        strip.text = element_text(color = "white"),
        plot.title.position = "plot",
        plot.title = element_text(color = "black", face = "bold")))
```

@@

```{r bigram-title}
bigram_tab(doe_db, Title) %>% 
  filter(n > 3) %>% 
  mutate(word = paste(word1, word2)) %>% 
  select(word, n) %>% 
  kbl(caption = "The bigram of the R-package titles as provided in the DESCRIPTION file in CRAN.", 
      col.names = c("Bigram", "Count"), 
      booktabs = TRUE, 
      linesep = "") 
```

```{r bigram-desc}
bigram_tab(doe_db, Description) %>% 
  filter(n > 4) %>% 
  mutate(word = paste(word1, word2)) %>% 
  select(word, n) %>% 
  kbl(caption = "The bigram of the R-package descriptions as provided in the DESCRIPTION file in CRAN.", 
      col.names = c("Bigram", "Count"), 
      booktabs = TRUE, 
      linesep = "") 
```

@@

```{r cran-data}
end <- Sys.Date() - 2 # usually 1-2 days data are not available yet
start <- end - years(2) + days(2)
dldat <- cran_downloads(doe_pkgs, from = start, to = end)
```

```{r download-hist, dependson = "cran-data"}
dldat %>% 
  group_by(package) %>% 
  summarise(total = sum(count)) %>%
  ggplot(aes(total)) + 
  geom_histogram(color = "white", fill = "#AD0059") + 
  scale_x_log10(label = comma) + 
  labs(x = glue("Total download counts from {start} to {end}"),
       y = "Number of packages") +
  scale_y_continuous(expand = c(0, 0))
```

```{r, dependson = "cran-data"}
dldat %>% 
  group_by(package) %>% 
  summarise(total = sum(count)) %>% 
  filter(total > 100000) %>% 
  pull(package)
```

```{r top5, dependson = "cran-data"}
ntop <- 10

@@

```{r download-barplot, dependson="cran-data"}
top5sum_df %>% 
  mutate(package = fct_reorder(package, total)) %>% 
  ggplot(aes(total, package)) +
  geom_col(aes(fill = package)) +
  labs(x = glue("Total downloads from {start} to {end}"),
       y = "Package") + 
  scale_x_continuous(labels = comma, expand = c(0, 0)) +
  scale_fill_discrete_qualitative(rev = TRUE) + 
  guides(fill = FALSE)
```

```{r pkg-updates, cache = TRUE, dependson=c("doe-cran", "top5")}
pkg_url <- "https://cran.r-project.org/web/packages/{pkg}/index.html"
pkg_archive <- "https://cran.r-project.org/src/contrib/Archive/{pkg}/"
pkg_updates <- map(top5, function(pkg) {
  last_update <- read_html(glue(pkg_url)) %>% 
    html_table() %>% 
    .[[1]] %>% 
    filter(X1=="Published:") %>% 
    pull(X2) %>% 
    ymd()

@@

```{r download-timeplot, dependson="top5"}
ggplot(top5_df, aes(date, count, color = package)) +
  # add shadow lines
  geom_line(data = rename(top5_df, package2 = package), 
            color = "gray", aes(group = package2)) +
  # add date when package was updated
  geom_vline(data = updates, aes(xintercept = update),
             linetype = "dashed", color = "#79003e") + 
  # the trend line
  geom_line() +
  scale_y_log10() +
  facet_grid(package ~ .) + 
  labs(title = glue("Top 5 downloaded DoE packages from {start} to {end}")) + 
  scale_color_discrete_qualitative() +
  guides(color = FALSE) 
```

### 2021-02-09 paper.tex
high level structure of the paper + some cosmetics for plots changed

\abstract{%
An abstract of less than 150 words.
}

@@

\caption{\label{tab:bigram-title}The bigram of the R-package titles as provided in the DESCRIPTION file in CRAN.}
\centering
\begin{tabular}[t]{l|r}
\hline
Bigram & Count\\
\hline
optimal design & 10\\
\hline
experimental design & 8\\
\hline
clinical trial & 5\\
\hline
dose finding & 5\\
\hline
sequential design & 5\\
\hline
block design & 4\\
\hline
microarray experiment & 4\\
\hline
response surface & 4\\
\hline
\end{tabular}
\end{table}

@@

\caption{\label{tab:bigram-desc}The bigram of the R-package descriptions as provided in the DESCRIPTION file in CRAN.}
\centering
\begin{tabular}[t]{lr}
\toprule
Bigram & Count\\
\midrule
experimental design & 11\\
optimal design & 10\\
package provide & 7\\
response surface & 7\\
factorial design & 6\\
\addlinespace
graphical user & 6\\
user interface & 6\\
block design & 5\\
contour plot & 5\\
design based & 5\\
\addlinespace
effect model & 5\\
fractional factorial & 5\\
microarray experiment & 5\\
mixed effect & 5\\
provide function & 5\\
\addlinespace
sample size & 5\\
sequential design & 5\\
\bottomrule
\end{tabular}
\end{table}

@@

\begin{Schunk}
\begin{Soutput}
#> `stat_bin()` using `bins = 30`. Pick better value with `binwidth`.
\end{Soutput}

@@

@@

\hypertarget{introduction}{%
\subsection{Introduction}\label{introduction}}

>>>>

\abstract{%
The critical role of data collection is well captured in the expression
``garbage in, garbage out'\,' -- in other words, if the collected data
is rubbish then no analysis, however complex it may be, can make
something out of it. The gold standard for data collection is through
well-designed experiments. Re-running an experiment is generally
expensive and in some cases difficult due to limited resources, The
design of experiment is important in getting a good value out of the
collected data. In this article I describe the current state of the
R-packages for the design of experiments through textual analysis and
download trends. I discuss also the software design of widely utilised R
packages in the field and conclude with discussion of some future
prospects for the field.
}

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

\hypertarget{explorative-analysis}{%
\section{Explorative Analysis}\label{explorative-analysis}}

\hypertarget{what-are-the-common-types-of-experimental-designs}{%
\subsection{What are the common types of experimental
designs?}\label{what-are-the-common-types-of-experimental-designs}}

\hypertarget{how-do-packages-interplay-with-each-other}{%
\subsection{How do packages interplay with each
other?}\label{how-do-packages-interplay-with-each-other}}

\hypertarget{which-packages-are-widely-utilised}{%
\subsection{Which packages are widely
utilised?}\label{which-packages-are-widely-utilised}}

\hypertarget{software-design}{%
\section{Software Design}\label{software-design}}

\hypertarget{propsects-and-discussion}{%
\section{Propsects and Discussion}\label{propsects-and-discussion}}

\newpage

@@

\caption{\label{tab:bigram-title}The bigram of the R-package titles as provided in the DESCRIPTION file in CRAN.}
\centering
\begin{tabular}[t]{lr}
\toprule
Bigram & Count\\
\midrule
optimal design & 10\\
experimental design & 8\\
clinical trial & 5\\
dose finding & 5\\
sequential design & 5\\
block design & 4\\
microarray experiment & 4\\
response surface & 4\\
\bottomrule
\end{tabular}
\end{table}

@@

\caption{\label{tab:bigram-desc}The bigram of the R-package descriptions as provided in the DESCRIPTION file in CRAN.}
\centering
\begin{tabular}[t]{lr}
\toprule
Bigram & Count\\
\midrule
experimental design & 11\\
optimal design & 10\\
package provide & 7\\
response surface & 7\\
factorial design & 6\\
graphical user & 6\\
user interface & 6\\
block design & 5\\
contour plot & 5\\
design based & 5\\
effect model & 5\\
fractional factorial & 5\\
microarray experiment & 5\\
mixed effect & 5\\
provide function & 5\\
sample size & 5\\
sequential design & 5\\
\bottomrule
\end{tabular}
\end{table}

@@

@@

\begin{Schunk}
\begin{Soutput}
#>  [1] "agricolae"   "AlgDesign"   "conf.design" "DiceDesign"  "DiceKriging"
#>  [6] "DoE.base"    "ez"          "FrF2"        "lhs"         "rsm"
\end{Soutput}
\end{Schunk}

@@

\hypertarget{introduction-1}{%
\subsection{Introduction}\label{introduction-1}}
